A scenario for a Planckian standard kilogram

### 2017-10-22 manuscript_gravitationalconstant.tex
Conclusion added, references added

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C].

@@

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature in themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some pression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

@@

A useful way to look at the samples of the posterior returned by the MCMC is doing a corner plot using a Python module named \textit{corner}\cite{corner}. The corner plot shows all the one and two dimensional projections of the posterior probability distributions and demonstrates quickly the covariances between parameters. Along the diagonal, the marginalized distribution for each parameter independently is shown.
In Fig.~\ref{fig:plotCorner4} an excerpt of the full corner plot is shown. Here we only depict the parameters $x_0$, $v_{y_0}$, $\omega_z$, $C_{xx}$ and $k$. These are the more important ones, because for example the main contribution to $\vec{\omega}$ comes from its $z$-component.\\

@@

To arrive at a best value for $k$ and an uncertainty, we take the median and the percentiles corresponding to one sigma from the distribution of the parameter $k$. For this example we get:
\begin{equation}\label{eq:kfit}
k = \left( 8.0088006 \, \substack{+0.0000005 \\ -0.0000008}\, \right) \cdot 10^{-10}
\end{equation}
Now we divide $k$ by the total mass of the two balls. Note, that we assume, that the uncertainty on the mass of the balls does not contribute to a larger uncertainty on $G$, because their mass should be determinable to a relative uncertainty of about $10^{-8}$, whereas the one of $k$ is roughly $10^{-7}$. We change the way to write the uncertainty from the 0.16 and 0.84 percentiles to the usual standard deviation, while adding some margin. With units put back in, we arrive at the following value for the gravitational constant $G$:
\begin{equation}\label{Gfit}
G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
\end{equation}
This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.\\

\section{Conclusion}

@@

Appendices are for material that is needed for completeness but
not sufficiently interesting to include in the main body of the paper.  Most
articles don't need any appendices, but feel free to use them when
appropriate.  This sample article needs an appendix only to illustrate how 
to create an appendix.

\begin{acknowledgments}

Need acknowledgments?
\end{acknowledgments}

@@

\bibitem{vanRavenzwaaij16} D. van Ravenzwaji, and P. Cassey, and S. D. Brown, ``A simple introduction to Markov Chain Monte-Carlo sampling,'' 
Psychonomic Bulletin {\&} Review, 1--12 (2016).  
% Note that the issue number (10) in this citation is required, because
% each issue of Physics Today starts over with page 1.  Also note the use of
% an en-dash (--), not a hyphen (-), for the page range.

@@

\bibitem{Tomilin99} D. K. Tomilin, ``Fine-structure constant and dimension analysis,'' 
European Journal of Physics (20.5), L39 (1999).

\bibitem{Fischer16} J. Ullrich, and J. Fischer, ``The new system of units,'' 
Nature Physics (12.1), 4--7 (2016).

\bibitem{Quinn14} T. Quinn, C. Speake, H. Parks and R. Davis, ``The BIPM measurements of the Newtonian constant of gravitation $G$'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (September 2014)

\bibitem{Newman14} R. Newman, M. Bantel, E. Berg and W. Cross, ``A measurement of $G$ with a cryogenic torsion pendulum'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (September 2014)

\bibitem{Luo09} J. Luo, W. Liu, L. Tu, C. Shao, L.Liu, S. Yang, Q. Li and Y. Zhang, ``Determination of the Newtonian gravitational constant $G$ with time-of-swing method'', \textit{Phys. Rev. Lett.}, 102:240801 (June 2009)

\bibitem{Schlamminger06} S. Schlamminger, E. Holzschuh, W. Kündig, F. Nolting, R.E. Pixley, J. Schnurr and U. Straumann, ``Measurement of Newton's gravitational constant", \textit{Phys. Rev. D}, 74:082001 (October 2009)

\bibitem{Parks10} H.V. Parks and J.E. Faller, ``Simple pendulum determination of the gravitational constant", \textit{Phys. Rev. Lett.}, 105:110801 (September 2010)

\bibitem{Rosi14} G. Rosi, F. Sorrentino, L. Cacciapuoti, M. Prevedelli and G.M. Tino, ``Precision measurement of the Newtonian gravitational constant using cold atoms", \textit{Nature}, 510:518-521 (June 2014)

@@

@@

>>>>

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C]\cite{Chao15}.

@@

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature in themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some expression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

@@

@@

A useful way to look at the samples of the posterior returned by the MCMC is doing a corner plot using a Python module named \textit{corner}\cite{corner}. The corner plot shows all the one and two dimensional projections of the posterior probability distributions and demonstrates quickly the covariances between parameters. Along the diagonal, the marginalized distribution for each parameter independently is shown.
In Fig.~\ref{fig:plotCorner4} an excerpt of the full corner plot is shown. \\
%
%Here we only depict the parameters $x_0$, $v_{y_0}$, $\omega_z$, $C_{xx}$ and $k$. These are the more important ones, because for example the main contribution to $\vec{\omega}$ comes from its $z$-component.\\
%

To arrive at a best value with uncertainty for the gravitational constant, we take the mean and standard deviation of the marginalized distribution of the parameter $k$ and divide by the total mass of the balls. In this example the relative uncertainty of $k$ is roughly $10^{-7}$, so if the masses are measured to about the same precision, we can write for the gravitational constant $G$:
\begin{equation}\label{Gfit}
G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
\end{equation}
This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.\\

%
%we take the median and the percentiles corresponding to one sigma from the distribution of the parameter $k$. For this example we get:
%\begin{equation}\label{eq:kfit}
%k = \left( 8.0088006 \, \substack{+0.0000005 \\ -0.0000008}\, \right) \cdot 10^{-10}
%\end{equation}
%Now we divide $k$ by the total mass of the two balls. Note, that we assume, that the uncertainty on the mass of the balls does not contribute to a larger uncertainty on $G$, because their mass should be determinable to a relative uncertainty of about $10^{-8}$, whereas the one of $k$ is roughly $10^{-7}$. We change the way to write the uncertainty from the 0.16 and 0.84 percentiles to the usual standard deviation, while adding some margin. With units put back in, we arrive at the following value for the gravitational constant $G$:
%\begin{equation}\label{Gfit}
%G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
%\end{equation}
%This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.\\
%

\section{Conclusion}
We have proposed a new way to measure the gravitational constant $G$ and estimated on what order of magnitude different disturbing effects occur and if one has to correct for them. We then simulated measurement data and showed how to extract the gravitational constant. Based on our estimations, it should be possible to measure the gravitational constant to about $10^{-7}$ relative uncertainty. This is a about a factor of 300 better than today's recommended CODATA value.\\
If we are a bit optimistic, a further increase in precision is not out of reach, so that about the same relative uncertainty as for Planck's constant $h$ ($1.2 \cdot 10^{-8}$)\cite{CODATA16} could be achieved. This then would allow to redefine the kilogram as an integer multiple of the Planck mass.\\

@@

%Appendices are for material that is needed for completeness but
%not sufficiently interesting to include in the main body of the paper.  Most
%articles don't need any appendices, but feel free to use them when
%appropriate.  This sample article needs an appendix only to illustrate how 
%to create an appendix.

\begin{acknowledgments}
Need acknowledgments?
\end{acknowledgments}

@@

\bibitem{Chao15} L.S. Chao, S. Schlamminger, D.B. Newell, J.R. Pratt, F. Seifert, X. Zhang, G. Sineriz, M. Liu and D. Haddad, ``A LEGO Watt balance: An apparatus to determine a mass based on the new SI", \textit{American Journal of Physics}, 83:11:913--922 (August 2015).

@@

\bibitem{Quinn14} T. Quinn, C. Speake, H. Parks and R. Davis, ``The BIPM measurements of the Newtonian constant of gravitation $G$'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (September 2014).

\bibitem{Newman14} R. Newman, M. Bantel, E. Berg and W. Cross, ``A measurement of $G$ with a cryogenic torsion pendulum'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (September 2014).

\bibitem{Luo09} J. Luo, W. Liu, L. Tu, C. Shao, L.Liu, S. Yang, Q. Li and Y. Zhang, ``Determination of the Newtonian gravitational constant $G$ with time-of-swing method'', \textit{Phys. Rev. Lett.}, 102:240801 (June 2009).

\bibitem{Schlamminger06} S. Schlamminger, E. Holzschuh, W. Kündig, F. Nolting, R.E. Pixley, J. Schnurr and U. Straumann, ``Measurement of Newton's gravitational constant", \textit{Phys. Rev. D}, 74:082001 (October 2009).

\bibitem{Parks10} H.V. Parks and J.E. Faller, ``Simple pendulum determination of the gravitational constant", \textit{Phys. Rev. Lett.}, 105:110801 (September 2010).

\bibitem{Rosi14} G. Rosi, F. Sorrentino, L. Cacciapuoti, M. Prevedelli and G.M. Tino, ``Precision measurement of the Newtonian gravitational constant using cold atoms", \textit{Nature}, 510:518-521 (June 2014).

@@

\bibitem{vanRavenzwaaij16} D. van Ravenzwaji, and P. Cassey, and S. D. Brown, ``A simple introduction to Markov Chain Monte-Carlo sampling,'' 
Psychonomic Bulletin {\&} Review, 1--12 (2016).

@@

\bibitem{CODATA16} P.J. Mohr, D.B. Newell and B.N. Taylor, ``CODATA recommended values of the fundamental physical constants: 2014*, \textit{Reviews of Modern Physics}, 88(3):035009 (July 2016).
